# PACE atopic dermatitis study

```{r setoptions, eval = TRUE, echo = FALSE}
opts_chunk$set(echo = FALSE, 
            warning = FALSE, 
            message = FALSE, 
            fig.width = 12, 
            fig.height = 7,
            cache = FALSE,
            fig.align = 'center')
```

```{r call_source}
read_chunk("~/PACE/pace_ad/reports/pace_ad_report.r")
```

``` {r load_data, results = "hide", message = FALSE, warning = FALSE}
```

## Aim:
To examine the association between DNA methylation and atopic dermatitis (AD) in early life


## Methods
* __Sample__: Danish individuals of a random subset of the GOYA cohort (N = `r nrow(goya)`). AD data was gathered at 7 years of age.
* __Tissue__: Cord blood (450k data)
* __Outcome definitions__:
	+ Childhood AD = A doctor has at some point said that the child has eczema/allergic rash, and this rash has been active within the last twelve months __OR__ The child has had an itchy rash within the last twelve months, and this rash has affected one of the following places: the folds of the elbows, behind the knees, in front of the ankles, under the buttocks, or around the neck, ears or eyes __OR__ The child has had either doctor diagnosed AD or rash as described in (2), but not active within the last 12 months.
	+ Early AD = A doctor has diagnosed the child with eczema/allergic rash that occurred before 2 years of age
	+ Persistant AD = A doctor has diagnosed the child with eczema/allergic rash before 2 years of age, and this rash has been active within the last twelve months
* __Models__:
	+ A. AD ~ methylation + sex + SVs

	+ B. AD ~ methylation + sex + SVs + maternal age + maternal smoking in pregnancy + maternal SES + gestational age

	+ C. AD ~ methylation + sex + SVs + maternal age + maternal smoking in pregnancy + maternal SES + gestational age + cell count (bakulski)


### Summary of data
```{r summary_table_setup, results = "hide", message = FALSE, warning = FALSE, echo = FALSE}
```
```{r summary_table, results = "asis", message = FALSE, warning = FALSE, echo = FALSE, fig.cap = cov_tab_cap}
```

## Results
```{r results_setup, results = "hide", message = FALSE, warning = FALSE, echo = FALSE}
```

```{r results_table1, results = "asis", message = FALSE, warning = FALSE, echo = FALSE, fig.cap = res_tab_cap}
```

```{r manhattan, results = "asis", message = FALSE, warning = FALSE, echo = FALSE, fig.cap = manhattans_cap}
```

