---
author: Thomas Battram
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: united
    number_sections: false
params:
  data: ""
  set_title: ""
  pheno: ""
title: "`r gsub('_', ' ', params$set_title)`"
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}

## pkgs
suppressWarnings(suppressPackageStartupMessages({
library(bookdown) # extra markdown options
library(knitr) # ...
library(kableExtra) # extra nice tables
library(tidyverse) # tidy code and data
library(ewaff) # For easy EWAS plots
library(cowplot) # plotting figures on a grid
library(RColorBrewer) # for plotting colours
library(usefunc) # own package of useful functions
}))

knitr::opts_chunk$set(echo = FALSE, 
           			  warning = FALSE, 
            		  message = TRUE, 
            		  cache = FALSE, 
            		  dpi = 300 
            		  )
```

```{r general-setup}
get_model <- function(res_file)
{
  stringr::str_extract(res_file, "m[1-3][a-c]")
}

get_definition <- function(phenotype)
{
  phenos <- list(childhood = "cases: AD by age 10, controls:no AD by age 10", 
                 early_onset = "cases: AD by age 2, controls: no AD by age 2", 
                 persistent = "AD by age 2 and persisting to ~ 8-10yrs, controls: no AD by age 10")
  return(phenos[[phenotype]])
}

# params <- list(data = '/newhome/tb13101/PACE/pace_ad/qc/report/input-files/m1a.RData /newhome/tb13101/PACE/pace_ad/qc/report/input-files/m1b.RData /newhome/tb13101/PACE/pace_ad/qc/report/input-files/m1c.RData', samplesizes = '/newhome/tb13101/PACE/pace_ad/qc/data/samplesizes.RData', set_title = 'Quality control report for PACE EWAS analysis of childhood atopic dermatitis', pheno = 'childhood')
data_files <- unlist(str_split(params$data, " "))
pheno <- params$pheno

data <- lapply(data_files, function(x) {out <- new_load(x); return(out)})
names(data) <- map_chr(data_files, get_model)
a_data <- data[[grep("a", names(data))]]
b_data <- data[[grep("b", names(data))]]
c_data <- data[[grep("c", names(data))]]

change_cell_colour <- function(df) 
{
    df[] <- lapply(df, function(x) {cell_spec(x, color = ifelse(trimws(x) == "TRUE", "red", "black"))}) 
    return(df)
}

make_table <- function(dat)
{
    dat <- dat %>%
      tidy_nums %>%
      tidy_colnames
    dat <- change_cell_colour(dat)
    kbl(dat, escape = FALSE) %>% 
        kable_styling() %>% 
        # pack_rows(index = c("model a" = 10, "model b" = 10, "model c" = 10)) %>% 
        scroll_box(width = "100%", height = "450px")
}
```

## Summary of analysis

This study is examining the association between DNA methylation and childhood eczema (up to 10 years). There are 3 definitions of eczema used throughout this study. This report is specific to EWAS of `r pheno` AD (`r get_definition(pheno)`).

For the QC, the checks laid out by the developers of [QCEWAS](https://pubmed.ncbi.nlm.nih.gov/28119308/) were followed (although the package itself was not used). To examine inflation of results the [ewaff](https://github.com/perishky/ewaff) package was used to generate lambdas and QQ-plots and the [BACON](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-016-1131-9) package to estimate inflation and bias factors.

## Model a

Logistic regression model:

\begin{equation}
  AD = meth + sex\ (+ selection)\ (+ ancestry)\ (+ batch)
  (\#eq:modela)
\end{equation}

where covariates in brackets are applied where appropriate.

### Samplesizes

```{r samplesizes-a}
make_table(a_data$samplesizes)
```

### Expected value checks

```{r stats-a}
make_table(a_data$stats)
```

### Inflation statistics

```{r inflation-a}
make_table(a_data$inflation)
```

### QQ

```{r qq-plots-a, results = "asis", out.width="100%"}
include_graphics(a_data$plots$qq)
```

### Beta

```{r beta-plots-a, results = "asis", out.width="100%"}
include_graphics(a_data$plots$beta_box)
```

### SE

```{r se-plots-a, results = "asis", out.width="100%"}
include_graphics(a_data$plots$se_box)
```

### Precision plot

```{r prec-plots-a, results = "asis", out.width="100%"}
include_graphics(a_data$plots$prec)
```

## Model b

Logistic regression model:

\begin{equation}
  AD = meth + sex + maternal age + maternal smoking + maternal SES + gestational age\ (+ selection)\ (+ ancestry)\ (+ batch)
  (\#eq:modelb)
\end{equation}

where covariates in brackets are applied where appropriate.

### Samplesizes

```{r samplesizes-b}
make_table(b_data$samplesizes)
```

### Expected value checks

```{r stats-b}
make_table(b_data$stats)
```

### Inflation statistics

```{r inflation-b}
make_table(b_data$inflation)
```

### QQ

```{r qq-plots-b, results = "asis", out.width="100%"}
include_graphics(b_data$plots$qq)
```

### Beta

```{r beta-plots-b, results = "asis", out.width="100%"}
include_graphics(b_data$plots$beta_box)
```

### SE

```{r se-plots-b, results = "asis", out.width="100%"}
include_graphics(b_data$plots$se_box)
```

### Precision plot

```{r prec-plots-b, results = "asis", out.width="100%"}
include_graphics(b_data$plots$prec)
```

## Model c

Logistic regression model:

\begin{equation}
  AD = meth + sex + maternal age + maternal smoking + maternal SES + gestational age + cell composition\ (+ selection)\ (+ ancestry)\ (+ batch)
  (\#eq:modelc)
\end{equation}

where covariates in brackets are applied where appropriate.

### Samplesizes

```{r samplesizes-c}
make_table(c_data$samplesizes)
```

### Expected value checks

```{r stats-c}
make_table(c_data$stats)
```

### Inflation statistics

```{r inflation-c}
make_table(c_data$inflation)
```

### QQ

```{r qq-plots-c, results = "asis", out.width="100%"}
include_graphics(c_data$plots$qq)
```

### Beta

```{r beta-plots-c, results = "asis", out.width="100%"}
include_graphics(c_data$plots$beta_box)
```

### SE

```{r se-plots-c, results = "asis", out.width="100%"}
include_graphics(c_data$plots$se_box)
```

### Precision plot

```{r prec-plots-c, results = "asis", out.width="100%"}
include_graphics(c_data$plots$prec)
```